---
title: "FY26 NIH Awarded Grants: Funding Intelligence Report"
author: "Oh, Eun Jeong"
date: "2026-05-23"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
    toc_depth: 3
    mathjax: default
runtime: shiny
---

<style type="text/css">
div.main-container {
  max-width: 1300px;
  margin: auto;
}

/* Insight boxes */
.insight-box {
  background: #eef4fb;
  color: #1a1a2e;
  border-left: 5px solid #0b6fa1;
  border-radius: 6px;
  padding: 15px 20px;
  margin: 16px 0;
  font-size: 0.96em;
  line-height: 1.7;
}

.insight-box strong {
  color: #0b6fa1;
  font-size: 0.92em;
}

.insight-box ul {
  margin: 8px 0 0 0;
  padding-left: 18px;
}

/* Strategy tip boxes */
.strategy-box {
  background: #fff8e7;
  border-left: 5px solid #f5a623;
  border-radius: 6px;
  padding: 14px 20px;
  margin: 14px 0;
  font-size: 0.95em;
}

.strategy-box .label {
  font-weight: 700;
  color: #c47d00;
  font-size: 0.85em;
  letter-spacing: 0.05em;
  text-transform: uppercase;
  margin-bottom: 5px;
}

/* Section intro text */
.section-intro {
  color: #555;
  font-size: 0.97em;
  line-height: 1.75;
  margin-bottom: 16px;
  border-bottom: 1px solid #e2e2e2;
  padding-bottom: 12px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 11)
```

# Overview

This page presents an R workflow for analyzing FY26 NIH-awarded grant data beyond simple word
counts. The analyses include agency-level funding summaries, a keyword-funding bubble map,
and TF-IDF-based agency fingerprints. The goal is to surface
actionable insights for researchers navigating the NIH funding landscape.

# Data Pre-processing

The data includes all NIH-awarded grants in the United States for FY26 up to the date specified
above.

```{r load-packages}
library(wordcloud)     # word cloud visualization
library(tm)            # text mining and corpus preprocessing
library(dplyr)         # data manipulation and summarisation
library(tidytext)      # tidy text mining: tokenization, tf-idf, stop words
library(ggplot2)       # static data visualization
library(pander)        # markdown table rendering
library(tidyr)         # data reshaping
library(stringr)       # string manipulation
library(plotly)        # interactive html charts
library(ggrepel)       # non-overlapping text labels on scatter plots
library(scales)        # axis label formatting (comma, percent, dollar)
library(shiny)         # reactive selectInput + renderPlotly for TF-IDF section
```

```{r load-data}
wd       = getwd()
filename = paste0(wd, "/SearchResult_Export_23May2026_012709.csv")
data     = read.csv(filename, skip = 7)

# helper: strip parenthetical phrases and digits from project titles
clean_title = function(x) {
  x %>%
    str_remove_all("\\s*\\([^\\)]+\\)") %>%
    str_remove_all("[0-9]+")
}

# agency-level summary: grant counts, total and mean award
df = data %>%
  rename(agency = Administering.IC) %>%
  group_by(agency) %>%
  summarise(
    total.number = n(),
    total.cost   = sum(Total.Cost, na.rm = TRUE),
    mean.cost    = mean(Total.Cost, na.rm = TRUE),
    median.cost  = median(Total.Cost, na.rm = TRUE)
  ) %>%
  mutate(
    total.cost.dollars = label_number(accuracy = 0.1,  scale_cut = cut_short_scale())(total.cost),
    mean.cost.dollars  = label_number(accuracy = 0.01, scale_cut = cut_short_scale())(mean.cost)
  ) %>%
  arrange(desc(total.number))

# token-level table: one row per (grant × word), retaining Total.Cost for later joins
token_df = data %>%
  rename(agency = Administering.IC) %>%
  mutate(title_clean = clean_title(Project.Title)) %>%
  unnest_tokens(word, title_clean) %>%
  anti_join(stop_words, by = "word") %>%
  filter(
    str_length(word) > 2,            # drop very short tokens
    !str_detect(word, "^[0-9]+$")    # drop numeric-only tokens
  )
```

---

# Agency Overview {.tabset}

## Total Funding

```{r plot-total-cost, fig.height=7}
ggplot(df) +
  geom_bar(
    aes(y = reorder(agency, total.cost), x = total.cost),
    stat = "identity", fill = "#16213e", width = 0.7
  ) +
  geom_text(
    aes(y = reorder(agency, total.cost), x = total.cost, label = total.cost.dollars),
    hjust = -0.2, colour = "#0b6fa1", size = 3.2
  ) +
  scale_x_continuous(
    labels = label_number(scale_cut = cut_short_scale()),
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title = "Total NIH Funding by Agency — FY26",
    x     = "Total Cost (USD) \n",
    y     = "\n Agency"
  ) +
  theme_minimal() +
  theme(
    plot.title   = element_text(face = "bold", colour = "darkblue", size = 14, hjust = 0.5),
    axis.title.x = element_text(face = "bold", colour = "darkblue", size = 12, vjust = -1.5),
    axis.title.y = element_text(face = "bold", colour = "darkblue", size = 12, vjust = 6),
    axis.text.y  = element_text(size = 9),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank()
  )
```

## Number of Grants

```{r plot-n-grants, fig.height=7}
ggplot(df) +
  geom_bar(
    aes(y = reorder(agency, total.number), x = total.number),
    stat = "identity", fill = "#0f3460", width = 0.7
  ) +
  geom_text(
    aes(y = reorder(agency, total.number), x = total.number, label = comma(total.number)),
    hjust = -0.2, colour = "#0b6fa1", size = 3.2
  ) +
  scale_x_continuous(
    labels = comma,
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title = "Number of Funded Grants by Agency — FY26",
    x     = "Number of Grants \n",
    y     = "\n Agency"
  ) +
  theme_minimal() +
  theme(
    plot.title   = element_text(face = "bold", colour = "darkblue", size = 14, hjust = 0.5),
    axis.title.x = element_text(face = "bold", colour = "darkblue", size = 12, vjust = -1.5),
    axis.title.y = element_text(face = "bold", colour = "darkblue", size = 12, vjust = 6),
    axis.text.y  = element_text(size = 9),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank()
  )
```

## Average Award Size

```{r plot-mean-award, fig.height=7}
ggplot(df) +
  geom_bar(
    aes(y = reorder(agency, mean.cost), x = mean.cost),
    stat = "identity", fill = "#533483", width = 0.7
  ) +
  geom_text(
    aes(y = reorder(agency, mean.cost), x = mean.cost, label = mean.cost.dollars),
    hjust = -0.2, colour = "#0b6fa1", size = 3.2
  ) +
  scale_x_continuous(
    labels = label_number(scale_cut = cut_short_scale()),
    expand = expansion(mult = c(0, 0.18))
  ) +
  labs(
    title    = "Average Award Size by Agency — FY26",
    subtitle = "Agencies with fewer but larger grants stand out here",
    x        = "Mean Award (USD) \n",
    y        = "\n Agency"
  ) +
  theme_minimal() +
  theme(
    plot.title    = element_text(face = "bold", colour = "darkblue", size = 14, hjust = 0.5),
    plot.subtitle = element_text(size = 10, colour = "grey40", hjust = 0.5),
    axis.title.x  = element_text(face = "bold", colour = "darkblue", size = 12, vjust = -1.5),
    axis.title.y  = element_text(face = "bold", colour = "darkblue", size = 12, vjust = 6),
    axis.text.y   = element_text(size = 9),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank()
  )
```

---

# Frequent Words in Project Titles

```{r wordcloud, fig.height=4.5}
# build a text corpus from cleaned project titles
title_corpus = data$Project.Title %>%
  clean_title() %>%
  VectorSource() %>%
  Corpus() %>%
  tm_map(removePunctuation) %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removeNumbers) %>%
  tm_map(stripWhitespace) %>%
  tm_map(removeWords, stopwords("english"))

par(bg = "black")
wordcloud(
  title_corpus,
  scale     = c(2, 1),
  min.freq  = 150,
  colors    = brewer.pal(7, "YlOrRd")
)
```

---

# Keywords × Funding: What Words Are Worth

<div class="section-intro">
Beyond word frequency, the critical question for grant writers is: <strong>which keywords appear
in higher-funded projects?</strong> For each keyword appearing in ≥ 50 project titles, the
analysis below computes (1) how many grants contain it and (2) the mean total award of those
grants. Plotting these two dimensions together reveals the "golden keywords" — terms that are
both common and associated with larger awards.
</div>

<div class="insight-box">
<strong>KEY INSIGHT</strong><br>
High-frequency keywords are not always high-value. Look for words in the
<em>upper-right quadrant</em> of the bubble chart — those that are <em>both</em> common
<em>and</em> linked to large awards. Bubble size adds a third dimension: the total NIH dollars
tied to that keyword across all grants.
</div>

```{r keyword-funding, echo=FALSE, fig.height=7, fig.width=11}
# compute per-keyword grant count and mean award
keyword_stats = token_df %>%
  group_by(word) %>%
  summarise(
    n_grants   = n_distinct(Application.ID),
    mean_cost  = mean(Total.Cost, na.rm = TRUE),
    total_cost = sum(Total.Cost,  na.rm = TRUE),
    .groups    = "drop"
  ) %>%
  filter(n_grants >= 50) %>%
  arrange(desc(mean_cost))

# bubble chart: frequency × mean award, sized by total funding
p_bubble = ggplot(
  keyword_stats,
  aes(x = n_grants, y = mean_cost, size = total_cost, label = word)
) +
  geom_point(aes(fill = mean_cost), shape = 21,
             colour = "white", alpha = 0.82, stroke = 0.4) +
  geom_text_repel(
    data      = keyword_stats %>%
      filter(
        n_grants  > quantile(n_grants,  0.75) |
        mean_cost > quantile(mean_cost, 0.75)
      ),
    aes(label = word),
    size = 2.8, colour = "#1a1a2e", fontface = "bold",
    box.padding = 0.4, max.overlaps = 25
  ) +
  scale_fill_gradient(
    low    = "#90caf9",
    high   = "#e94560",
    labels = label_number(scale_cut = cut_short_scale())
  ) +
  scale_size_continuous(
    range  = c(2, 14),
    breaks = function(x) pretty(x, n = 4),
    labels = label_number(accuracy = 0.1, scale_cut = cut_short_scale())
  ) +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  guides(
    size = guide_legend(
      override.aes = list(shape = 21, fill = "grey60", colour = "white", stroke = 0.4)
    ),
    fill = guide_colorbar()
  ) +
  labs(
    title    = "Keyword Portfolio Map — FY26 NIH Grants",
    subtitle = "X: grant count containing keyword  |  Y: mean award of those grants  |  Bubble size: total $ tied to keyword",
    x        = "\n Number of Grants Containing Keyword",
    y        = "Mean Award Size (USD) \n",
    fill     = "Mean Award",
    size     = "Total Funding"
  ) +
  theme_minimal() +
  theme(
    plot.title    = element_text(face = "bold", colour = "darkblue", size = 13, hjust = 0.5),
    plot.subtitle = element_text(size = 9, colour = "grey40", hjust = 0.5),
    axis.title.x  = element_text(face = "bold", colour = "darkblue", size = 11, vjust = -1.5),
    axis.title.y  = element_text(face = "bold", colour = "darkblue", size = 11, vjust = 3),
    legend.position = "right"
  )

p_bubble
```

```{r keyword-funding-table, echo=FALSE}
# top 20 keywords ranked by mean award size
keyword_stats %>%
  arrange(desc(mean_cost)) %>%
  head(20) %>%
  mutate(
    mean_cost  = label_number(accuracy = 0.01, scale_cut = cut_short_scale())(mean_cost),
    total_cost = label_number(accuracy = 0.1,  scale_cut = cut_short_scale())(total_cost)
  ) %>%
  rename(
    Keyword         = word,
    `# Grants`      = n_grants,
    `Mean Award`    = mean_cost,
    `Total Funding` = total_cost
  ) %>%
  knitr::kable(caption = "Top 20 Keywords by Mean Award Size (minimum 50 grants)")
```

<div class="strategy-box">
<div class="label">📌 Grant Writing Strategy</div>
Keywords near the top of this table appear in titles of grants with the highest average awards.
If your science genuinely involves these concepts, make them visible — in your title, specific
aims, and abstract. NIH study sections recognize these as signals of ambitious, large-scope
projects.
</div>

---

# TF-IDF: What Makes Each IC Unique

<div class="section-intro">
Simple word frequency tells you what is common across all of NIH. <strong>TF-IDF</strong>
(Term Frequency – Inverse Document Frequency) tells you what is <em>distinctively
characteristic</em> of each agency. The score combines two components:

$$\text{TF-IDF}(t,\, d) \;=\; \frac{\text{count of term } t \text{ in agency } d}{\text{total terms in agency } d} \;\times\; \log\!\left(\frac{\text{total number of agencies}}{\text{number of agencies containing } t}\right)$$

- **TF (Term Frequency):** how often a word appears within a given agency's grant titles,
  relative to all words that agency uses. A word used heavily by NIDA gets a high TF for NIDA.
- **IDF (Inverse Document Frequency):** the log of the ratio of total agencies to the number
  of agencies that use the word at all. Words that appear in nearly every IC — such as
  *disease*, *clinical*, or *study* — receive an IDF close to zero, which collapses their
  TF-IDF score regardless of local frequency. Words concentrated in only one or two agencies
  receive a high IDF and therefore a high TF-IDF score.

Multiplying TF × IDF rewards words that are both locally frequent *and* globally rare across
ICs. In practice this surfaces the "fingerprint" vocabulary of each agency — terms like
*addiction* for NIDA or *vision* for NEI. A proposal whose core terminology aligns with an
IC's TF-IDF fingerprint signals to reviewers that the science belongs there.

*Note: because each agency panel is scaled independently below, bar lengths reflect within-agency
rank rather than absolute cross-agency comparisons. Agencies with large, diverse portfolios
(e.g., NINDS, NHLBI) naturally produce lower absolute TF-IDF values because no single term
dominates — this is expected, not a data quality issue. TF-IDF also measures cross-agency
distinctiveness rather than a direct comparison against the NIH-wide average rate; for the
latter, a weighted log-odds approach via the* `tidylo` *package would be more appropriate.*
</div>

<div class="insight-box">
<strong>KEY INSIGHT</strong><br>
If your proposal's core keywords have high TF-IDF in a specific agency, that IC is likely
your best-fit home. A mismatch between your scientific vocabulary and the target IC's fingerprint
is a signal to either reconsider the agency or reframe your proposal language.
</div>

```{r tfidf-compute, echo=FALSE}
# compute TF-IDF treating each agency as a "document"
agency_tfidf = token_df %>%
  count(agency, word, sort = TRUE) %>%
  bind_tf_idf(word, agency, n) %>%
  arrange(agency, desc(tf_idf))
```

```{r tfidf-shiny, echo=FALSE}
# all agencies ordered by grant count
agencies_ordered = df %>%
  arrange(desc(total.number)) %>%
  pull(agency)

# build per-agency top-12 data once
tfidf_top12 = lapply(agencies_ordered, function(ag) {
  agency_tfidf %>%
    filter(agency == ag) %>%
    slice_max(tf_idf, n = 12) %>%
    arrange(tf_idf) %>%
    mutate(agency_label = ag)
}) %>% bind_rows()

# native HTML select — scrollable, easy to use
selectInput(
  inputId  = "selected_agency",
  label    = "Select Agency:",
  choices  = agencies_ordered,
  selected = agencies_ordered[1],
  width    = "200px"
)

renderPlotly({
  ag  = input$selected_agency
  dat = tfidf_top12 %>% filter(agency_label == ag)

  plot_ly(
    dat,
    x             = ~tf_idf,
    y             = ~factor(word, levels = word),  # preserve sorted order
    type          = "bar",
    orientation   = "h",
    marker        = list(color = "#0b6fa1"),
    hovertemplate = "<b>%{y}</b><br>TF-IDF: %{x:.5f}<extra></extra>"
  ) %>%
    layout(
      title = list(
        text = paste0(ag, ": Top 12 Characteristic Keywords"),
        font = list(size = 13, color = "darkblue")
      ),
      xaxis = list(
        title     = "TF-IDF Score (higher = more distinctive to that agency)",
        titlefont = list(size = 10)
      ),
      yaxis = list(
        title      = "",
        tickfont   = list(size = 10),
        automargin = TRUE
      ),
      margin = list(l = 10, t = 50)
    )
})
```

<div class="strategy-box">
<div class="label">📌 Grant Writing Strategy</div>
Use this chart to verify your agency choice. Find the IC you are targeting and check whether
your proposal's core concepts match its keyword fingerprint. If your key terms do not appear
here, your agency selection may be off — or you have an opportunity to pioneer a new direction,
but be prepared to justify it to reviewers.
</div>

---

# Funding Intelligence Summary

<div class="insight-box">
<strong>TAKEAWAYS FOR GRANT WRITERS</strong>
<ul>
  <li><strong>Golden keywords</strong> (high frequency AND high mean award) are the most
  strategic terms to incorporate authentically into your title and specific aims.</li>
  <li><strong>TF-IDF fingerprints</strong> reveal each IC's scientific identity. Words with high
  TF-IDF are rare across NIH but concentrated in one agency — matching your vocabulary to the
  IC's fingerprint signals reviewer fit.</li>
  <li>A mismatch between your scientific vocabulary and the target IC's TF-IDF profile is a
  signal to either reconsider the agency or reframe your proposal language.</li>
</ul>
</div>

# Contact

Questions, suggestions, and collaboration inquiries are welcome. Please feel free to contact me
if you are interested in adapting this workflow to other funding datasets or research settings.

**Eun Jeong Oh, PhD**  
Assistant Professor of Biostatistics, Research Intelligence  
Northwell Health  
GitHub: https://github.com/oheunj  
Email: eoh2 [at] northwell [dot] edu

**Code Availability**  
The data access and preparation code is shown above for transparency. The code used to generate
the figures and tables in this report is not displayed in the rendered page, but it is available
upon request.

---

© 2026 Eun Jeong Oh. All rights reserved.
